---
title: 'Day 6: Linear models cont.'
author: "J. Lacasa"
date: "10/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Packages  

## Review  

```{r data}
usda <- read.csv("../data/USDAsurvey_00-18.csv")
legume <- read.csv("../data/legume.csv")
```

## Exploratory data analysis  

## Linear Mixed Effects  

**Fixed effects**  
Effect ON the variable. (not affecting the error, just the expected Y value).   

**Random effects**  
* NOT AFFECTING the expected Y value.  
* Effect on the “variability” of the variable (Sigma).  
* Most basic component is the RESIDUALS (e) given by replications/repetitions.  
* Also BLOCKS.  

***  
```{r Expected marginal means}
options('contrasts')
options(contrasts = c("contr.sum", "contr.poly")) # set the contrasts
options('contrasts')

usda_mixed <- lmer(Value ~ State + (1|fYear), data = usda)

usda_mixed_means <- emmeans(usda_mixed, #~State|Year,
                      specs = pairwise ~ State)

```

***  

Random effects are iid ~ N(0,var1)  
```{r QQ plot for random effects}
randef <- ranef(usda_mixed)[[1]] 

randef %>% 
  ggplot(aes(sample=`(Intercept)`))+
  stat_qq()+
  stat_qq_line()
```

***  

```{r pairwise comp}
#Set up a compact letter display of all pair-wise comparisons
cld(usda_mixed_means,
    adjust = "none",
    Letters = letters,
    reversed = TRUE,
    comparisons = TRUE) -> usda_mixed_cld

# default plot 

# ggplot -> necessary steps ?
```

***  

```{r mixedeff}
usda_mixed_cld <- usda_mixed_cld %>% as.data.frame() 

usda_mixed_cld %>%
  ggplot(aes(emmean, State))+
  geom_point()+
  geom_text(aes(label = .group), nudge_y = 0, nudge_x = 3)

usda_mixed_cld$.group <- gsub(" ", "", usda_mixed_cld$.group, fixed = TRUE)

usda_mixed_cld <- usda_mixed_cld %>%
  arrange(emmean) %>% 
  rownames_to_column("order") %>% 
  mutate(order = as.numeric(order))

usda_mixed_cld <- usda_mixed_cld %>% mutate(State = factor(State, levels = usda_mixed_cld$State[usda_mixed_cld$order]))
# ggplot 
```

***  

```{r better mixedeff viz}
usda_mixed_cld %>%
  ggplot(aes(emmean, State))+
  geom_errorbarh(aes(xmin = asymp.LCL, xmax = asymp.UCL), color = "grey70", height = .2)+
  geom_point()+
  geom_text(aes(label = .group, x = emmean + nchar(.group)*1.2), nudge_y = 0.1)+
  labs(x = expression(Mean~Yield~(Bu~acre^{-1})))+
  theme_bw()+
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()) -> usda_mixed_ggplot; usda_mixed_ggplot 
```



# References  

* **Model building:** Check out Chapter 10 in *Hobbs, N. T., and Hooten, M. B. (2015). Bayesian models: a statistical primer for ecologists. Princeton University Press.*  

* [Dr. Leo Bastos 2019 R workshop](https://www.youtube.com/watch?v=8xyi9By3fM4)  
